EDIpack2: interoperable Lanczos-based solver for generic quantum impurity problems

### 2025-03-29 edipack2.tex
Updated draft. Up to Bath parametrization. Still a lot to write ....

\subsection{Conserved quantum numbers}\label{sSecQNs}
In order to circumvent the exponential scaling of the dimensions of
the problem scales it is necessary to take into account suitable symmetries.
Indeed, considering operators ${\cal Q}$ such that $[H, {\cal Q}]=0$
reduces of the Fock space into several symmetry sectors
labelled by given quantum numbers $\vec{Q}$. 
In the context of quantum impurity problems two symmetries are often considered: i)
conservation of the total occupation $N$ and ii) the
conservation of the total magnetization $S_z$. Although the total spin
operator $S^2$ can also be conserved, the difficult  implementation
and the marginal gain makes it often non convenient to include this symmetry.

In \NAME we consider three different cases:
\begin{itemize}
\item{\bf normal:}  $\vec{Q}=[N,S_z]\equiv[N_\up,N_\dw]$
\item{\bf superc:}  $\vec{Q}=S_z$
\item{\bf nonsu2:} $\vec{Q}=N$
\end{itemize}

The {\bf normal} case consider independent conservation of the total occupation $N$
and the total magnetization $S_z$ or, equivalently, the total number of electrons with spin up $N_\up$ and down
$N_\dw$. Optionally, we consider the case in which the symmetry
applies separately for each orbital and spin,
i.e. $\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This special case has been
extensively discussed in \onlinecite{Amaricci2022} so it won't be covered in
this work.

The {\bf superc} case deals with the conservation of the total
magnetization only, so that total charge may not be conserved. This
case includes a description of $s$-wave superconductivity, featuring
intra- and inter-orbital components.

Finally, the {\bf nonsu2} case consider the conservation of the total
charge, whereas the spin symmetry group is not fully
conserved. Although there many possible realization of this scenario,
this particularly applies to the presence of local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, the emergence of in-plane spin ordering~\cite{KM} or
in-plane spin-triplet exciton condensation~\cite{ExcitonPRB,Amaricci,Blason}.

From a computational point of view the construction of a symmetry
sector corresponds to the determination of a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ relating the states $\ket{i}$
belonging to the sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
Fock space. Operatively, the map corresponds to an integer rank-1
array of dimension $D_\SS$ which is the {\it dimension} of the
sector. According to the previously introduced symmetries we
have:
\begin{itemize}
\item{\bf normal:}  $D_\SS = \binom{N_s}{N_\up}\binom{N_s}{N_\dw}$ 
\item{\bf superc:}  $D_\SS = \sum_m 2^{N_s-S_z-2m}\binom{N_s}{N_s-S_z-2m}\binom{S_z+2m}{m}$
\item{\bf nonsu2:} $D_\SS=\binom{2N_s}{N}$
\end{itemize}

@@

The symmetry sector reduction of the Fock space induces a block
digonal form to the Hamiltonian matrix. Each block, labelled by a
value of the quantum numbers $\vec{Q}$ has dimension
$D_{\SS(\vec{Q}}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix. In the {\bf
  normal} case $H_\SS$ takes a particularly symmetric form thanks to
the product structure of the sector and its map~\cite{amaricci2022}.
The analysis of the spectrum is then reconducted to the inspection of
the Hamiltonian in each symmetry sector. Should particular constraint
holds the search can be limited to only particular sector, further
reducing the computational cost.

@@

\subsubsection{GFmatrix}
One of the main goal of the code is to evaluate dynamical correlations
functions $\ibra \TT_\pm[ A(t)
A^+ ]\iket$. As we shall see in the following, using Krylov method it is
possible to express the dynamical correlations in terms of a suitably
truncated Kallen-Lehmann spectral sum of the form
$\tfrac{1}{Z}\sum_n e^{-\beta E_n} \sum_{m=1}^{N} \tfrac{|w_{mn}|^2}{ z
  - dE_{mn}}\equiv $ where $w_{mn}$ is a weight set by the 
projection of the $m^{\rm th}$ eigenstate onto the $n^{\rm th}$
component of the Krylov basis (which reduce the sector Hamiltonian
into a partial tri-diagonal form) while $dE_{mn}=E_m-E_n$ is an excitation energy. 
The {\tt ED\_GFMATRIX} module contains a class efficiently storing all
the {\it weights} $w_{mn}$ and {\it poles} $dE_{mn}$ contributing to
a specific dynamical correlation functions. Specifically, the {\tt
  gfmatrix} object implements a dynamical multi-layer data structure
gatering weights and poles from any initial state and for any
combination of operators contributing to the dynamical correlation
function.

The use of this class enables the istantan

@@

\subsection{Bath parametrization}\label{sSecBath}

>>>>

\subsection{Conserved quantum numbers}\label{sSecQNs}
In order to circumvent the exponential scaling of the dimensions of
the problem scales it is necessary to take into account suitable symmetries.
Indeed, considering operators ${\cal Q}$ such that $[H, {\cal Q}]=0$
reduces of the Fock space into several symmetry sectors
labelled by given quantum numbers $\vec{Q}$. 
In the context of quantum impurity problems two symmetries are often considered: i)
conservation of the total occupation $N$ and ii) the
conservation of the total magnetization $S_z$. Although the total spin
operator $S^2$ can also be conserved, the difficult  implementation
and the marginal gain makes it often non convenient to include this symmetry.
In \NAME we consider three different cases which are controlled by the
input variable {\tt ed\_mode}={\bf normal}, {\bf superc}, {\bf
  nonsu2}.

The {\bf normal} case deals with independent conservation of the total occupation $N$
and the total magnetization $S_z$ or, equivalently, the total number of electrons with spin up $N_\up$ and down
$N_\dw$. Optionally, we consider the case in which the symmetry
applies separately for each orbital and spin,
i.e. $\vec{N_\s}=[N^1_\s,\dots,N^{N_\a}_\s]$. This special case has been
extensively discussed in \onlinecite{Amaricci2022} so it won't be covered in
this work.
%
The {\bf superc} case deals with the conservation of the total
magnetization only, so that total charge may not be conserved. This
case includes a description of $s$-wave superconductivity, featuring
intra- and inter-orbital components.
%
Finally, the {\bf nonsu2} case consider the conservation of the total
charge, whereas the spin symmetry group is not fully
conserved. Although there many possible realization of this scenario,
this particularly applies to the presence of local spin-orbit coupling
$\vec{L}\cdot\vec{S}$~\cite{something}, the emergence of in-plane spin ordering~\cite{KM} or
in-plane spin-triplet exciton condensation~\cite{ExcitonPRB,Amaricci,Blason}.  
%
From a computational point of view the construction of a symmetry
sector corresponds to the determination of a injective map
$\MM:\SS_{\vec{Q}}\rightarrow \FF$ relating the states $\ket{i}$
belonging to the sector $\SS_{\vec{Q}}$ to the states $\ket{I}$ in the
Fock space. Operatively, the map corresponds to an integer rank-1
array of dimension $D_\SS$ which is the {\it dimension} of the
sector. 
The table~\ref{TabSector} summarizes the properties of the symmetry
sectors.

\begin{table}%[ht]
  \label{TabSector}
\begin{center}
\begin{tabularx}{\linewidth}{ |X|X|X| } 
 \hline
  {\tt ed\_mode} & {\it Quantum Numbers} & {\it Sector Dimension} \\
  \hline
  {\bf normal} & $[N,S_z]\equiv[N_\up,N_\dw]$ &
                                                $\binom{N_s}{N_\up}\binom{N_s}{N_\dw}$
  \\
  \hline
  {\bf superc} & $S_z\equiv N_\up-N_\dw$ &  $\sum_m 2^{N_s-S_z-2m}\binom{N_s}{N_s-S_z-2m}\binom{S_z+2m}{m}$
  \\
  \hline
  {\bf nonsu2} & $N \equiv[N_\up+N_\dw$ & $\binom{2N_s}{N}$ \\ 
 \hline
\end{tabularx}
\end{center}
\caption{Ciao}
\end{table}

@@

The presence of a symmetry induces a factorization of the Fock space,
in turn inducing a block diagonal form to the Hamiltonian matrix.
Each block, labelled by the quantum numbers $\vec{Q}$, has dimension
$D_{\SS(\vec{Q})}$. The sector Hamiltonian matrix $H_\SS$ is represented in the
basis $\ket{i}\in\SS_{\vec{Q}}$ as a sparse matrix. In the {\bf
  normal} case $H_\SS$ takes a particularly symmetric form thanks to
the product structure of the sector and its map~\cite{amaricci2022}.
The analysis of the spectrum is then reconducted to the inspection of
the Hamiltonian in each symmetry sector. Should particular constraint
holds the search can be limited to only particular sector, further
reducing the computational cost.

@@

\subsubsection{GFmatrix}
One of the main goal of the code is to evaluate dynamical correlations
functions (DCF) $\ibra \TT[ A(t) A^+ ]\iket$. As we shall see in the following, using Krylov method it is
possible to express the dynamical correlations in terms of a suitably
truncated Kallen-Lehmann spectral sum of the form
$\tfrac{1}{Z}\sum_n e^{-\beta E_n} \sum_{m=1}^{N} \tfrac{|w_{mn}|^2}{ z
  - dE_{mn}}$ where $w_{mn}$ is a weight set by the 
projection of the $m^{\rm th}$ eigenstate onto the $n^{\rm th}$
component of the Krylov basis (which reduce the sector Hamiltonian
into a partial tri-diagonal form) while $dE_{mn}=E_m-E_n$ is an excitation energy. 
The {\tt ED\_GFMATRIX} module contains a class which efficiently store all
the {\it weights} $w_{mn}$ and {\it poles} $dE_{mn}$ contributing to
a specific correlation function. Specifically, the {\tt
  gfmatrix} object implements a dynamical multi-layer data structure
storing any DCF as the set of all the weights and poles,
from any contributing eigenstate  and for any
combination of operators appearing in its definition.
The use of this class enables the istantantaneous evaluation of a
given  DCF  $\GG(z)$ for any $z\in\CCC$ the complex frequency plane
using a suitable compressed form.

\subsection{Bath parametrization}\label{sSecBath}
The bath parametrization is a key feature in determining the properties
of the quantum impurity problem.
Following the structure of Eq.~\ref{Hbath} the bath can be
parametrized by two terms: the Hamiltonian matrices $h^p$ and the
amplitudes $V^p$, for $p=1,\dots,N_{bath}$.
Internally, the bath is represented by a dedicated object {\tt
  effective\_bath} defined in {\tt ED\_VARS\_GLOBAL}. 
On the
user side all parameters are packed into a rank-1 array of
doubles handled using reverse communication strategy.
This ensures that a local array is always conserved by the user, while
it prevents direct access to the internal copy.

The bath topology, i.e. the links between the $N_b$ electronic levels
assigned to the bath, is determined by the input variable {\tt
  bath\_type} among 4 different choices: {\bf normal}, {\bf hybrid},
{\bf replica} and {\bf general} (see Fig.\ref{figBaths}).  
%
\paragraph{{\tt bath\_type}={\bf normal}} The bath is formed out of $N_{bath}$
electronic levels coupled to each of the impurity orbital. The total
count of bath levels is $N_b = N_\a N_{bath}$. The bath Hamiltonan
matrices include a parametrization diagonal in the orbital and spin space: 
$h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.
If {\tt ed\_mode}={\bf superc} one needs to account for anomalous
amplitudes connecting bath levels with opposite spins. So we consider
an additional set of parameters, diagonal in orbital space but
off-diagonal in the spins:
$h^p_{\a\b\s\s'}=\Delta^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}}$.
This choice corresponds to consider, for every orbital component,
bath matrices with the following structure in the Nambu space:
$\hat{h}^p_{\a\b} = \e_{\a}\delta_{\a\b}\tau_0 +
\Delta_{\a}\delta_{\a\b}\tau_x$. 
The hybridization amplitudes between the impurity and the bath levels
include, for any value of {\tt ed\_mode} a set of parameters diagonal in both
spin and orbital space: $V^p_{\a\b\s\s'}=V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.
If {\tt ed\_mode}={\bf nonsu2} the an additional set including terms
describing spin-flip processes (as total magnetization is not
conserved) should be included $V^p_{\a\b\s\s'} =
W^p_{\a}\delta_{\a\b}\delta_{\s \overline{\s}}$.

\paragraph{{\tt bath\_type}={\bf hybrid}} The bath is formed out of
$N_{bath}$ coupled to all the impurity levels, correspondingly
$N_b\equiv N_{bath}$ is the total count of electronic bath levels. 
The main parametrization of the bath Hamiltonian matrices is still
diagonal in both orbital and spin space, i.e.
$h^p_{\a\b\s\s'}=\varepsilon_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.
As for the previous case, the {\bf superc} mode requires the inclusion
of a further set of parameters taking care of the anomalous components:
$h^p_{\a\b\s\s'}=\Delta^p_{\a}\delta_{\a\b}\delta_{\s
  \overline{\s}}$.
The key difference with respect to the {\bf normal} case lies in the
off-diagonal nature of the hybridization amplitudes:
$V^p_{\a\b\s\s'}=V^p_{\a\b\s}\delta_{\s\s'}$. If total magnetization
is not conserved, i.e. for {\tt ed\_mode}={\bf nonsu2}, an additional
set of parameters should be included to describe spin-flip processes:
$V^p_{\a\b\s\s'} = W^p_{\a\b}\delta_{\s \overline{\s}}$. 
These sets of hybridization parameters allow to capture the effects of locally
hybridized impurity orbitals. This comes at the cost of a
slightly harder optimization process (see \secu{sSecFit}) specially
for systems with reduced number of available bath levels with respect
to the impurity.

\paragraph{{\tt bath\_type}={\bf replica}/{\bf general}} A more flexible
parametrization of the bath is represented by this topology. The
original idea of this approach is to give to each bath element a structure which {\it replicates}
that of the impurity while keeping a diagonal coupling between bath
elements and impurity. This offloads the difficulties related to the
representation and optimization of strucured quantum impurities to the
bath Hamiltonan rather than the hybridizations.
Taking a sligthly more general point of view we consider a user
defined matrix basis $\{ \Gamma^\nu \}_{\nu=1,\dots,N_{sym}}$ in the (Nambu-)spin-orbital
space and parametrize any bath Hamiltonian as:
$$
h^p_{\a\b\s\s'} = \sum_{\nu=1}^{N_{sym}} \lambda^p_\nu \Gamma^\nu_{\a\b\s\s'}
$$
where $\lambda^p_\nu\in\RRR$ are suitable variational parameters. 
The choice of the matrix basis can be inspired either by the internal structure
of the quantum impurity, i.e. $\hat{h}^0$, or be determined
case-by-case by the properties of the problem.
The total number of electronic levels used to describe the bath is in
this case: $N_b = N_{sym} N_{bath}$. 
For the {\bf replica}  topology the coupling between the impurity and
each bath elements is diagonal in the spin, orbital and internal bath
structure: $V^p_{\a\b\s\s'} = V^p$. The
{\bf general} setup introduces a generalization introducing a
dependence on the internal spin and orbital indices: $V^p_{\a\b\s\s'}
= V^p_{\a\s}\delta_{\a\b}\delta_{\s\s'}$.

All the procedures concerning the bath are grouped into a set of
modules wrapped by {\tt ED\_BATH}.

* get dimension

* replica/general bath setup

@@

\subsection{Bath Functions}\label{sSecFunc}
